
---
title: "Homework 2"
author: "Anna Ostropolets"
date: "09/29/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

## Problem 1
```{r trash_wheel, message=FALSE}
# Importing Mr. Trash Wheel dataset with omitting non-data entries and rows without dumpster-specific data, 
# cleaning column names, rounding sports balls and converting them to integers.
trash_data = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = "A2:N370") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls)) %>% 
  mutate(sports_balls = as.integer(sports_balls))

```

```{r precipitation,message=FALSE}
# Importing precipitation data for 2017 and 2018 years omitting the rows without precipitation data. 
# Also adding year
precipitation_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation", 
    range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2017) 

precipitation_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation", 
    range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2018)

# combining two sets together with a helper dataframe
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precipitation_17_and_18 =
bind_rows(precipitation_2017,precipitation_2018)
left_join(precipitation_17_and_18, month_df, by = "month")

```

Mr. Trash Wheel dataset consists of `r trash_data %>% count() %>% pull(n)` records about `r trash_data %>% pull(dumpster) %>% last()` dumpsters. It describes collection of trash in different years (from `r trash_data %>% pull(date) %>% first()` to `r trash_data %>% pull(date) %>% last()`). It describes how much trash in tons and in cubic yards was collected by each dumpster at a given date with various types of trash: bottles, cigarette butts, grocery  bags,  etc. In total, Mr. Trash Wheel collected `r trash_data %>% pull(weight_tons) %>% sum()` tons of trash, which is an equivalent of `r trash_data %>% pull(volume_cubic_yards) %>% sum()` cubic yards of trash. The median number of sports balls pulled from dumpsters in 2017 was `r trash_data %>% filter(year==2017) %>% pull(sports_balls) %>% median()`. 
Precipitation dataset contains the information about `r precipitation_17_and_18 %>% count() %>% pull(n)` precipitation in 2017 and 2018, with `r precipitation_17_and_18 %>% filter(year==2017) %>% pull(total) %>% sum()` total precipitation in 2017 and `r precipitation_17_and_18 %>% filter(year==2018) %>% pull(total) %>% sum()` total precipitation in 2018. 


## Problem 2

```{r transit_data, message=FALSE}
# Importing and cleaning transit data
transit_data = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character())) %>% 
  janitor::clean_names() %>% 
  select( line, station_name, station_latitude, station_longitude, route1:route11, entrance_type ,entry, vending, ada) %>% 
  mutate( entry = recode (entry, "YES" = TRUE,"NO" = FALSE)) %>%
  mutate( vending = recode (vending, "YES" = TRUE,"NO" = FALSE))

```

NYC transit dataset consists of `r transit_data %>% count() %>% pull(n)` records about `r transit_data %>% distinct(station_name) %>% count()  %>% pull (n)` stations. It contains information about station latitude and longitude, routes served, entrance and exit info etc.
To clean it up, I selected only needed variables, converted YES/NO variables into logical variables and..
The dimensions of the dataset are `r transit_data %>% dim_desc()`.
The dataset isn't tidy.

There are `r transit_data %>% distinct(line,station_name) %>% count()  %>% pull (n)` stations. 
`r filter(transit_data, ada == "TRUE") %>% distinct(station_name, line) %>% count()  %>% pull (n)` stations are ADA compliant. The proportion of station entrances without vending but allowing entrance is `r (filter(transit_data, vending == "FALSE" & entry == "TRUE") %>% count() %>% pull(n)) / (filter(transit_data, vending == "FALSE") %>% count() %>% pull(n))`.

```{r transit_data_cleaned, message=FALSE}
 
# Reformatting transit data so that route number and route name are distinct variables.
transit_cleaned =
  transit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",#cut column namesv
    values_to = "route_line") %>% 
  relocate("route_line") %>% 
  drop_na(route_line)

```
There are `r filter(transit_cleaned, route_line == "A") %>% distinct(station_name, line) %>% count() %>% pull (n)` distinct stations that serve the A train. Out of them, `r filter(transit_cleaned, route_line == "A" & ada == "TRUE") %>% distinct(station_name, line) %>% count() %>% pull (n)` are ada compliant.

## Problem 3

```{r pols, message=FALSE}
# Loading in FiveThirtyEight data step by step
#Read in pols-month.csv and clean
pols = 
  read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = if_else(prez_gop==1, "gop", "dem")) %>% 
  select(-prez_gop, -prez_dem, -day) %>% 
  arrange(year, month) %>% 
  select(year, month, everything())
```

```{r snp, message=FALSE}
#Read in snp.csv and clean
snp = 
  read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day) %>% 
  arrange(year, month) %>% 
  select(year, month, everything())
```

```{r unemployment, message=FALSE}
#Read in unemployment.csv and tiday to long format with same variable names/formats
unemployment = 
  read_csv("./data/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment_rate") %>% 
  mutate(month = month.name[match(month, month.abb)]) %>% 
  janitor::clean_names() %>% 
  arrange(year, month)
```

```{r full_data, message=FALSE}
#Merge snp into pols then merge unemployment into result
full_data = 
  full_join(pols, snp, by = c("year", "month")) %>% 
  full_join(unemployment, by = c("year", "month"))

```
FiveThirtyEight dataset contains data on pols, unemployment and snp.Pols describes the positions held in congress 
 from `r pols %>% pull(month) %>% first()` `r pols %>% pull(year) %>% first()` to `r pols %>% pull(month) %>% last()` `r pols %>% pull(year) %>% last()`. 
Unemployment dataset has the unemployment rate from `r unemployment %>% pull(month) %>% first()` `r unemployment %>% pull(year) %>% first()` to `r unemployment %>% pull(month) %>% last()` `r unemployment %>% pull(year) %>% last()`. 
SnP has the closing values for trading SP index from `r snp %>% pull(month) %>% first()` `r snp %>% pull(year) %>% first()` to `r snp %>% pull(month) %>% last()` `r snp %>% pull(year) %>% last()`.  
The final dataset has the following variables: `r full_data %>% tbl_vars()`.
Dimensions are `r full_data %>% dim_desc()` and the dates are from `r full_data %>% pull(month) %>% first()` `r full_data%>% pull(year) %>% first()` to `r full_data %>% pull(month) %>% last()` `r full_data %>% pull(year) %>% last()`.



