---
title: "Homework 2"
author: "Anna Ostropolets"
date: "09/29/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

## Problem 1
```{r}
# Importing Mr. Trash Wheel dataset with omitting non-data entries and rows without dumpster-specific data, 
# cleaning column names, rounding sports balls and converting them to integers.
trash_data = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = "A2:N370") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls)) %>% 
  mutate(sports_balls = as.integer(sports_balls))

```

```{r}
# Importing precipitation data for 2017 and 2018 years omitting the rows without precipitation data. 
# Also adding year
precipitation_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation", 
    range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    mutate(year = 2017) 

precipitation_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation", 
    range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    mutate(year = 2018)

# combining two sets together
precipitation_17_and_18 =
bind_rows(precipitation_2017, precipitation_2018) %>%
mutate(month = month.name[month])
```

Mr. Trash Wheel dataset consists of `r trash_data %>% count() %>% pull(n)` records about `r trash_data %>% pull(dumpster) %>% last()` dumpsters. It describes collection of trash in different years (from `r trash_data %>% pull(date) %>% first()` to `r trash_data %>% pull(date) %>% last()`). It describes how much trash in tons and in cubic yards was collected by each dumpster at a given date with various types of trash: bottles, cigarette butts, grocery  bags,  etc. In total, Mr. Trash Wheel collected `r trash_data %>% pull(weight_tons) %>% sum()` tons of trash, which is an equivalent of `r trash_data %>% pull(volume_cubic_yards) %>% sum()` cubic yards of trash. The median number of sports balls pulled from dumpsters in 2017 was `r trash_data %>% filter(year==2017) %>% pull(sports_balls) %>% median()`. 
Precipitation dataset contains the information about `r precipitation_17_and_18 %>% count() %>% pull(n)` precipitation in 2017 and 2018, with `r precipitation_17_and_18 %>% filter(year==2017) %>% pull(total) %>% sum()` total precipitation in 2017 and `r precipitation_17_and_18 %>% filter(year==2018) %>% pull(total) %>% sum()` total precipitation in 2018. 


